---
title: "Assignment 4 - Applying meta-analytic priors"
author: "Riccardo Fusaroli"
date: "3/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 4

In this assignment we do the following:
- we reproduce the meta-analysis of pitch SD from last semester in a Bayesian framework
- we reproduce the pitch SD in schizophrenia analysis from last semester using both a conservative and a meta-analytic prior
- we assess the difference in model quality and estimates using the two priors.

The questions you need to answer are: 
1. What are the consequences of using a meta-analytic prior? 
2. Evaluate the models with conservative and meta-analytic priors. Discuss the effects on estimates. Discuss the effects on model quality. 
3. Discuss the role that meta-analytic priors should have in scientific practice. Should we systematically use them? Do they have drawbacks? Should we use them to complement more conservative approaches? How does the use of meta-analytic priors you suggest reflect the skeptical and cumulative nature of science?

### Step by step suggestions

Step 1: Reproduce the meta-analysis of pitch sd from previous studies of voice in schizophrenia
- the data is available as Assignment4MetaData.xlsx
- Effect size (cohen's d), sd and variance are already calculated (you're welcome!)
- Since we're only interested in getting a meta-analytic effect size, let's take a shortcut and use bromance magic (brms): https://mvuorre.github.io/post/2016/2016-09-29-bayesian-meta-analysis/




```{r}
#parallel::detectCores()  # how many cores in computer

devtools::install_github("stan-dev/rstan", ref = "develop", subdir = "rstan/rstan", build_vignettes = FALSE, dependencies = TRUE)

devtools::install_github("mvuorre/brmstools")

```

```{r}
library(rethinking)
library(readxl)
library(brms)
library(metafor)
library(lme4)
library(tidyverse)
library(brmstools)

```



```{r}
# Read data
d_meta = read_excel("Assignment4MetaData.xlsx")

d_pitch = read_excel("Assignment4PitchDatav2.xlsx")

# Use only complete cases
dm <- d_meta[ complete.cases(d_meta$MeanES), ]

# Mean and SE tell us all we need to know to get the priors - what is the estimated effect (mean) and how uncertain it is (SE) – the more uncertain, the less effect it has, and vice versa



# Visualization of the estimated effect
ggplot(dm, aes(x=MeanES, y=StudyRef)) +
    geom_segment(aes(x = MeanES-SdES*2, xend = MeanES+SdES*2, y=StudyRef, yend=StudyRef)) +
    geom_point()


# What we did previously
m = rma(yi = MeanES, vi = VarianceES, sdi = SdES, data = dm, slab = StudyRef)
metafor::forest(m)
summary(m)


# In a baysian perspective
m1 <- brm(MeanES|se(SdES) ~ 1 + (1|StudyRef), 
          data = dm,
          family = gaussian(),
          prior = NULL,
          cores = 2, 
          chain = 2, 
          warmup = 2000,
          iter = 5000)
summary(m1)

# Forest plot
brmstools::forest(m1,
       show_data = TRUE,
       av_name = "Effect size") + geom_vline(aes(xintercept = 0), linetype = "dashed")
```



Step 2: Prepare the pitch SD data from last year
- the data is available as Assignment4PitchData.csv (thanks Celine)
- We do not know how to build random effects, yet. So we need to simplify the dataset to avoid interdependence between datapoint: How?
- Also, let's standardize the data, so that they are compatible with our meta-analytic prior (Cohen's d is measured in SDs).

```{r}
# Simplify the dataset to avoid interdependence between data points
dp <- d_pitch %>% 
  group_by(ID_unique) %>% 
  summarise_all(mean)

# Make df
dp = as.data.frame(dp)

# check if it matches
length(unique(d_pitch$ID_unique)) == nrow(dp) # shoud be TRUE

# Standardize
dp$PitchMean.s <- scale(dp$PitchMean)[,1]
dp$PitchSD.s <- scale(dp$PitchSD)[,1]
  # For the ungrouped analysis
d_pitch$PitchMean.s <- scale(d_pitch$PitchMean)[,1] 
d_pitch$PitchSD.s <- scale(d_pitch$PitchSD)[,1]

# Set variables
dp$diagnosis <- as.factor(dp$diagnosis)
dp$ID_unique <- as.factor(dp$ID_unique)
  # For the ungrouped analysis
d_pitch$diagnosis <- as.factor(d_pitch$diagnosis)
d_pitch$ID_unique <- as.factor(d_pitch$ID_unique)

```


Step 3: Build a regression model predicting Pitch SD from Diagnosis.
- how is the outcome distributed? (likelihood function)
- how are the parameters of the likelihood distribution distributed? Which predictors should they be conditioned on?
- use a skeptical/conservative prior for the effects of diagnosis. Remember you'll need to motivate it.
- Describe and plot the estimates. Evaluate model quality
```{r}
# Simple model without priors
m2 <- brm(PitchSD.s ~ diagnosis, 
          data = dp,
          family = gaussian(),
          prior = NULL,
          cores = 3, 
          chain = 3, 
          warmup = 2000, 
          iter = 5000)
summary(m2)

#temp_df <- posterior_samples(m2) 
#dens(temp_df$b_diagnosis1)
#stanplot(m2)
#plot(m2)


# Define weak priors          
m3_formula <- bf(PitchSD.s ~ diagnosis + (diagnosis|p|ID_unique), 
                 sigma ~ diagnosis + (diagnosis|p|ID_unique))
get_prior(m3_formula, d_pitch) # examine priors
prior <- c(prior(normal(0,1), class = Intercept), 
           prior(normal(0,0.5), class = b, coef = diagnosis1),   
           prior(normal(0,0.3), class = b, coef = diagnosis1, dpar = sigma),  
           prior(cauchy(0,1), class = Intercept, dpar = sigma) 
           )

  # verify that the priors found their way into Stan's model code
make_stancode(m3_formula,
              data = d_pitch, family = gaussian(),
              prior = prior)

# Model with weak priors
m3 <- brm(m3_formula, 
          data = d_pitch,
          family = gaussian(),
          prior = prior,
          cores = 3,     # have to use 1 chain and 3 cores, otherwise it takes too much time to run
          chain = 1, 
          warmup = 1000, 
          iter = 2000, 
          control = list(adapt_delta = 0.9))
summary(m3)

stanplot(m3)
plot(m3)

# Plots 
temp_df <- posterior_samples(m3)
dens(temp_df$b_diagnosis )
pp_check(m3, nsamples = 100, group = d_pitch$diagnosis)
?pp_check(m3, nsamples = 100)

test_data <- data.frame(diagnosis = c(rep(0, 1000),rep(1, 1000)), ID_unique = c(seq(3000,4999)))
test_data <- data.frame(diagnosis = c(rep(0, 1000),rep(1, 1000)))

#temp <- predict(m3, newdata = test_data, allow.new.levels = TRUE) #allow.new.levels = T does 

ggplot(temp_df1, aes(x = est, color = type)) + geom_density(adjust = 1) + labs(x = "Estimated")


# With conservative priors
prior.c <- c(prior(normal(0,1), class = Intercept), 
           prior(normal(0,0.1), class = b, coef = diagnosis1), 
           prior(normal(0,0.3), class = b, coef = diagnosis1, dpar = sigma),
           prior(cauchy(0,1), class = Intercept, dpar = sigma) 
           )

m3.c <- brm(m3_formula, 
          data = d_pitch,
          family = gaussian(),
          prior = prior.c,
          cores = 3, 
          chain = 1, 
          warmup = 1000, 
          iter = 2000, 
          control = list(adapt_delta = 0.9))
summary(m3.c)



# PitchSD ~ 1 + diagnosis + (1 + diagnosis|ID)
#m2.stan <- map2stan( 
 # alist(
  #  avgPitchSD_s ~ dnorm( mu , sigma ) ,
   # mu <- a + b*diagnosis ,
    #a ~ dnorm(0,1),                 # -3/+3
    #b ~ dnorm(-.6,.3),               
          # we expect the diff between 2 groups to be between -0.6 and +0.6
          # gaussian: all possible values are in the range equal to 3 SD on each side
          # 0.1 to be more sceptical
          # mean = -.6 - using mean(int) 
          # SD = .3 - using SE(int) as SD - estimating the underlying true phenomenon
    #sigma ~ dcauchy(0,2)   
    #) ,
  #data=dp)
#precis(m2.stan)

# Plots
#plot(m2.stan)
#pairs(m2.stan)

```



Step 4: Now re-run the model with the meta-analytic prior
- Describe and plot the estimates. Evaluate model quality
```{r}
# With meta-analytic prior
prior.m <- c(prior(normal(0,1), class = Intercept), 
           prior(normal(-0.54,0.23), class = b, coef = diagnosis1), 
           prior(normal(0,0.3), class = b, coef = diagnosis1, dpar = sigma),
           prior(cauchy(0,1), class = Intercept, dpar = sigma) 
           )


m3.m <- brm(m3_formula, 
          data = d_pitch,
          family = gaussian(),
          prior = prior.m,
          cores = 3, 
          chain = 1, 
          warmup = 1000, 
          iter = 2000, 
          control = list(adapt_delta = 0.9))
summary(m3.m)


```


Step 5: Compare the models
- Plot priors and posteriors of the diagnosis effect in both models
- Compare posteriors between the two models
- Compare their relative distance from truth (WAIC)
- Discuss how they compare and whether any of them is best.
```{r}
# Plot priors
x <- seq(-2,2, length=1e5)
y <- dnorm(x, 0, 0.5) #original
y.s <- dnorm(x, 0, 0.1) #sceptical
y.m <- dnorm(x, -0.54, 0.23) #meta
prior_df <- data.frame(x = rep(x,3), y = c(y, y.s, y.m), prior = c(rep("original", length(y)),
                                                                   rep("sceptical", length(y.s)),
                                                                   rep("meta", length(y.m))
                                                                   ))
ggplot(prior_df, aes(x = x, y = y, color = prior)) + geom_line() 

# Plot posterior
post_samples <- c(posterior_samples(m3)$b_diagnosis1, posterior_samples(m3.c)$b_diagnosis1, posterior_samples(m3.m)$b_diagnosis1)
post_df <- data.frame(post_samples = post_samples, model = c(rep("original", 1000),
                                                                   rep("sceptical", 1000),
                                                                   rep("meta", 1000)
                                                                   ))
ggplot(post_df, aes(x = post_samples, color = model)) + geom_density(adjust = 1) 

# Compare models
waic <- brms::WAIC(m3, m3.c, m3.m)
weights <- brms::model_weights(m3, m3.c, m3.m, weights = "waic")
waic
weights

# Plots - quality check
  # weak
pp_check(m3,nsamples =200)
  # sceptical
pp_check(m3.c,nsamples =200)
  # informative
pp_check(m3.m,nsamples =200)


```




Step 6: Prepare a nice write up of the analysis and answer the questions at the top.

Optional step 7: how skeptical should a prior be?
- Try different levels of skepticism and compare them using WAIC.

Optional step 8: Include other predictors
- Do age, gender and education improve the model?
- Should they be main effects or interactions?

Optional step 9: Bromance magic.
- explore the bromance code below including random effects (by default with weakly informative priors)
- learn how to change the prior
- explore effects of trial, age, gender, including the appropriate random slopes
- compare the models you created using WAIC and posterior predictive check (pp_check())


```{r}

brm_out <- brm(PitchSD ~ 1 + Diagnosis  +(1|ID_unique/Study), # Outcome as a function of the predictors as in lme4. 
               data=Data, # Define the data
               family=gaussian(), # Define the family. 
               iter = 5000, warmup = 2000, cores = 4)
summary(brm_out1)
plot(brm_out1)

```

